Supporting information: Tail associations and their impact on extinction risk

### 2019-10-28 SuppMat.Rmd
running with common spearman's rho

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
library(abind)
source("mtime.R")
# Note : for moderate tail dep. we used Kendall's tau = 0.6 to get C or SC
```

```{r get_ktau_common, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("ExtremeTailDep.R"))}
set.seed(seed)
library(copula)
source("./ExtremeTailDep.R")

np<-100000

#for(i in 1:5){ # for loop to ensure the consistency of Kendall's tau with specified np
  xtc_r<-retd(n=np,d=2,rl=1,mn=0,sdev=1) # copula with extreme right tail

  start_time <- Sys.time()
  ktau_common<- cor(xtc_r[,1],xtc_r[,2],method="kendall") #get the kendall's tau
  end_time <- Sys.time()

#  cat("points = ",np, ", ktau_common = ", ktau_common, ", run time = ", (end_time -start_time),"\n")
#}

# see the change in 3rd decimal place
#points =  1e+05 , ktau_common =  0.7512526 , run time =  1.837576 
#points =  1e+05 , ktau_common =  0.7507594 , run time =  1.864947 
#points =  1e+05 , ktau_common =  0.7497681 , run time =  1.862879 
#points =  1e+05 , ktau_common =  0.7487744 , run time =  1.854317 
#points =  1e+05 , ktau_common =  0.7516491 , run time =  1.840355

saveRDS(ktau_common,"./Results/ktau_common.RDS")

@@

# for clayton : left tail dep.
copC<-claytonCopula(3)
parC<-iTau(copC,tau=ktau_common)

@@

# calculate tau with sampling variation
kcor_el<-cor(xtc_l,xtc_l,method = "kendall")
kcor_el<-kcor_el[1,2]

kcor_er<-cor(xtc_r,xtc_r,method = "kendall")
kcor_er<-kcor_er[1,2]

kcor_ml<-cor(mtc_l,mtc_l,method = "kendall")
kcor_ml<-kcor_ml[1,2]

kcor_mr<-cor(mtc_r,mtc_r,method = "kendall")
kcor_mr<-kcor_mr[1,2]
```

@@

# extreme left tail
plot(xtc_l[1:500,1],xtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
#legend(x="topleft",legend="A",bty="n",cex=2.5)
text(x=0.1,y=0.9,"A",cex=2.5)
text(x=0.8,y=0.1,round(kcor_el,4),cex=2)

# moderate left tail
plot(mtc_l[1:500,1],mtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
text(x=0.1,y=0.9,"B",cex=2.5)
text(x=0.8,y=0.1,round(kcor_ml,4),cex=2)

# moderate right tail
plot(mtc_r[1:500,1],mtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
text(x=0.1,y=0.9,"C",cex=2.5)
text(x=0.8,y=0.1,round(kcor_mr,4),cex=2)

# extreme right tail
plot(xtc_r[1:500,1],xtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
text(x=0.1,y=0.9,"D",cex=2.5)
text(x=0.8,y=0.1,round(kcor_er,4),cex=2)

@@

```{r sim_linear_mtd, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("ext_risk_cop_age_str.R"))}
# extinction risk with moderate tail dep. for single patch linear age structure model

@@

numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf
ktau<-ktau_common  # kendall's tau

# for clayton
copC<-claytonCopula(3)
parC<-iTau(copC,tau=ktau)
par<-parC
fam<-3

@@

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_linear",sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

@@

# Now for survival clayton : should be the same parameter as of clayton if kendall's tau is the same

@@

```{r sim_nonlinear_mtd, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("ext_risk_cop_age_str.R"))}
# extinction risk with moderate tail dep. for single patch non-linear age structure model

@@

numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-100
ktau<-ktau_common  # kendall's tau

# for clayton
copC<-claytonCopula(3)
parC<-iTau(copC,tau=ktau)
par<-parC
fam<-3

@@

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_ktau_",ktau,"_nonlinear",sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

@@

```{r sim_ricker_scl_1, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

@@

numsims<-10000
numsteps<-25
numlocs<-5
K<-50
scl<-1
ktau<-ktau_common
ext_thrs<-K/10
model<-"ricker"

@@

for(i in c(1:length(rlist))){
    r<- rlist[i]
    params<-c(r,K)
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

for(r in rlist){
    #cat("r=",r,"\n")
    params<-c(r,K)
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_hassell_scl_1, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

@@

numsims<-10000
numsteps<-25
numlocs<-5
a<-0.5
b<-100
scl<-1
ktau<-ktau_common

@@

for(i in c(1:length(rlist))){
    r<- rlist[i]
    K_e<-((r^(1/b))-1)/a
    ext_thrs<-K_e/10
    params<-c(r,a,b)
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

# For global dispersion
res_hassell_global<-c()
for(r in rlist){
    #cat("r=",r,"\n")
    K_e<-((r^(1/b))-1)/a
    ext_thrs<-K_e/10
    params<-c(r,a,b)
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_msmith_scl_0.8, echo=F, cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

model<-"msmith"
scl<-0.8
ktau<-ktau_common

@@

for(i in c(1:length(rlist))){
    r<- rlist[i]
    params<-c(r,a,b)
    K_e<-((r-1)^(1/b))/a
    ext_thrs<-K_e/10
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

# For global dispersal in D matrix 
res_msmith_global<-c()
for(r in rlist){
    #cat("r=",r,"\n")
    params<-c(r,a,b)
    K_e<-((r-1)^(1/b))/a
    ext_thrs<-K_e/10
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_pennycuick_scl_1, echo=F, cache=T, cache.extra=list(seed,ktau_common,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

model<-"pennycuick"
scl<-1
ktau<-ktau_common

@@

for(i in c(1:length(rlist))){
    r<- rlist[i]
    params<-c(r,a,b)
    K_e<-(b/a)*(a+log(r-1))
    ext_thrs<-K_e/10
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

# For global dispersal in D matrix 
res_pennycuick_global<-c()
for(r in rlist){
    #cat("r=",r,"\n")
    params<-c(r,a,b)
    K_e<-(b/a)*(a+log(r-1))
    ext_thrs<-K_e/10
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_verhulst_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

model<-"verhulst"
scl<-0.6
ktau<-ktau_common

@@

    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
                 scl=scl,ktau=ktau,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_abrewer_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,ktau_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

model<-"austinbrewer"
scl<-0.6
ktau<-ktau_common

@@

The moderate tail association was modeled by using the 
concept of "copula", a statistical tool to model 
dependencies. Appendix \ref{SM-intro_copula} summarized
a brief introduction of "copulas" and their relation to tail associations, symmetries 
with examples of available copula families 
from `R` packages (`copula` and `VineCopula`). We used bivariate Clayton 
copula to model moderate left-tail and Survival Clayton copula to model 
moderate right-tail association in the vital rates. Both copulae had the same Kendall 
correlation ($\tau =$ `r ktau_common`) but opposite association in their respective tails. 
We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
right (Survival Clayton) tail associated bivariate copulas (see `BiCopSim` function from
`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
by number of simulations ($=10000$) we used. 
We chose parameters for each copula family using 
`iTau` function from `copula` package so that both copula had
same Kendall's correlation ($\tau =$ `r ktau_common`). Entries within each column of
matrix $M$, then became independent of each other but 
entries between columns had a specific tail association
with positive Kendall's correlation.

@@

For moderately left-tail associated 
noises we drew points from a $n$ dimensional 
Clayton copula with same Kendall correlation ($\tau =$ `r ktau_common`) between
any two out of $n$ patches and made 
the marginals normal with zero mean and standard deviation $\sigma$ 
(see the specified value $\sigma$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}). For moderately 
right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
from Clayton copula so that all the statistical properties remained 
same except for the direction of tail association. We generated
extreme left- or right-tail associated noises via algorithm stated in 
Appendix \ref{SM-extr_tail_dep}.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
library(abind)
source("mtime.R")
```

```{r get_srho_common, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("ExtremeTailDep.R"))}
set.seed(seed)
library(copula)
source("./ExtremeTailDep.R")

np<-1000000
srho_common<-0.875 #7/8

xtc_r<-retd(n=np,d=2,rl=1,mn=0,sdev=1) # copula with extreme right tail

saveRDS(srho_common,"./Results/srho_common.RDS")

@@

# for clayton : left tail dep.
copC<-claytonCopula(3)
parC<-iRho(copC,rho=srho_common)

@@

# calculate rho with sampling variation
scor_el<-cor(xtc_l,xtc_l,method = "spearman")
scor_el<-scor_el[1,2]

#scor_er<-cor(xtc_r,xtc_r,method = "spearman")
scor_er<- -scor_el # Note it should be opposite to scor_el

scor_ml<-cor(mtc_l,mtc_l,method = "spearman")
scor_ml<- scor_ml[1,2]

#scor_mr<-cor(mtc_r,mtc_r,method ="spearman")
scor_mr<- -scor_ml # Note it should be opposite to scor_ml
```

@@

# extreme left tail
plot(xtc_l[1:500,1],xtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
#legend(x="topleft",legend="A",bty="n",cex=2.5)
text(x=0.1,y=0.9,"A",cex=2.5)
text(x=0.8,y=0.1,round(scor_el,4),cex=2)

# moderate left tail
plot(mtc_l[1:500,1],mtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
text(x=0.1,y=0.9,"B",cex=2.5)
text(x=0.8,y=0.1,round(scor_ml,4),cex=2)

# moderate right tail
plot(mtc_r[1:500,1],mtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
text(x=0.1,y=0.9,"C",cex=2.5)
text(x=0.8,y=0.1,round(scor_mr,4),cex=2)

# extreme right tail
plot(xtc_r[1:500,1],xtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
text(x=0.1,y=0.9,"D",cex=2.5)
text(x=0.8,y=0.1,round(scor_er,4),cex=2)

@@

```{r sim_linear_mtd, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("ext_risk_cop_age_str.R"))}
# extinction risk with moderate tail dep. for single patch linear age structure model

@@

numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf
srho<-srho_common  # Spearman's rho

# for clayton
copC<-claytonCopula(3)
parC<-iRho(copC,rho=srho)
par<-parC
fam<-3

@@

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_scor_",srho,"_linear",sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

@@

# Now for survival clayton : should be the same parameter as of clayton if Spearman's rho is the same

@@

```{r sim_nonlinear_mtd, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("ext_risk_cop_age_str.R"))}
# extinction risk with moderate tail dep. for single patch non-linear age structure model

@@

numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-100
srho<-srho_common  # Spearman's rho

# for clayton
copC<-claytonCopula(3)
parC<-iRho(copC,rho=srho)
par<-parC
fam<-3

@@

tempo<-paste("./Results/age_str_results/",BiCopName(family=fam),"_par_",par,"_scor_",srho,"_nonlinear",sep="")
if (!dir.exists(tempo)){
  dir.create(tempo)
}

@@

```{r sim_ricker_scl_1, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

@@

numsims<-10000
numsteps<-25
numlocs<-5
K<-50
scl<-1
srho<-srho_common
ext_thrs<-K/10
model<-"ricker"

@@

for(i in c(1:length(rlist))){
    r<- rlist[i]
    params<-c(r,K)
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

for(r in rlist){
    #cat("r=",r,"\n")
    params<-c(r,K)
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_hassell_scl_1, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

@@

numsims<-10000
numsteps<-25
numlocs<-5
a<-0.5
b<-100
scl<-1
srho<-srho_common

@@

for(i in c(1:length(rlist))){
    r<- rlist[i]
    K_e<-((r^(1/b))-1)/a
    ext_thrs<-K_e/10
    params<-c(r,a,b)
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

# For global dispersion
res_hassell_global<-c()
for(r in rlist){
    #cat("r=",r,"\n")
    K_e<-((r^(1/b))-1)/a
    ext_thrs<-K_e/10
    params<-c(r,a,b)
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_msmith_scl_0.8, echo=F, cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

model<-"msmith"
scl<-0.8
srho<-srho_common

@@

for(i in c(1:length(rlist))){
    r<- rlist[i]
    params<-c(r,a,b)
    K_e<-((r-1)^(1/b))/a
    ext_thrs<-K_e/10
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

# For global dispersal in D matrix 
res_msmith_global<-c()
for(r in rlist){
    #cat("r=",r,"\n")
    params<-c(r,a,b)
    K_e<-((r-1)^(1/b))/a
    ext_thrs<-K_e/10
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_pennycuick_scl_1, echo=F, cache=T, cache.extra=list(seed,srho_common,a_pennycuick,b_pennycuick,rc_pennycuick,r_bf_pennycuick,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

model<-"pennycuick"
scl<-1
srho<-srho_common

@@

for(i in c(1:length(rlist))){
    r<- rlist[i]
    params<-c(r,a,b)
    K_e<-(b/a)*(a+log(r-1))
    ext_thrs<-K_e/10
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

# For global dispersal in D matrix 
res_pennycuick_global<-c()
for(r in rlist){
    #cat("r=",r,"\n")
    params<-c(r,a,b)
    K_e<-(b/a)*(a+log(r-1))
    ext_thrs<-K_e/10
    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=K_e,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_verhulst_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

model<-"verhulst"
scl<-0.6
srho<-srho_common

@@

    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=F,plotteron = F,resloc=resloc, getacf=F)

@@

    ss<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,p0=p0,params=params,ext_thrs=ext_thrs,
                 scl=scl,srho=srho,model=model,disp_everywhere=T,plotteron = F,resloc=resloc, getacf=F)

@@

```{r sim_abrewer_scl_0.6, echo=F,  cache=T, cache.extra=list(seed,srho_common,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("mydispersal.R")

model<-"austinbrewer"
scl<-0.6
srho<-srho_common

@@

The moderate tail association was modeled by using the 
concept of "copula", a statistical tool to model 
dependencies. Appendix \ref{SM-intro_copula} summarized
a brief introduction of "copulas" and their relation to tail associations, symmetries 
with examples of available copula families 
from `R` packages (`copula` and `VineCopula`). We used bivariate Clayton 
copula to model moderate left-tail and Survival Clayton copula to model 
moderate right-tail association in the vital rates. Both copulae had the same Kendall 
correlation ($\rho =$ `r srho_common`) but opposite association in their respective tails. 
We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
right (Survival Clayton) tail associated bivariate copulas (see `BiCopSim` function from
`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
by number of simulations ($=10000$) we used. 
We chose parameters for each copula family using 
`iRho` function from `copula` package so that both copula had
same Kendall's correlation ($\rho =$ `r srho_common`). Entries within each column of
matrix $M$, then became independent of each other but 
entries between columns had a specific tail association
with positive Kendall's correlation.

@@

For moderately left-tail associated 
noises we drew points from a $n$ dimensional 
Clayton copula with same Kendall correlation ($\rho =$ `r srho_common`) between
any two out of $n$ patches and made 
the marginals normal with zero mean and standard deviation $\sigma$ 
(see the specified value $\sigma$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}). For moderately 
right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
from Clayton copula so that all the statistical properties remained 
same except for the direction of tail association. We generated
extreme left- or right-tail associated noises via algorithm stated in 
Appendix \ref{SM-extr_tail_dep}.
